Black swans in ecological time series\\ Black swans in population dynamics\\ Heavy-tailed ecological time series\\ Black swans in ecological dynamics\\ Black swan ecological dynamics\\ \ldots

### 2014-11-18 text/supp.tex
Add to supp caption
Lots of edits and wrap text

\item We removed all populations with uneven sampling intervals, i.e.\ we removed populations that didn't have a constant difference between the ``decimal year begin'' and ``decimal year end'' columns.

@@

\item Finally, we removed populations with the following GPDD Main IDs, which we discovered data entry errors when verifying the populations with suspected black swans: \texttt{1207} because the 1957 data point was entered as 2 but should have been 27 \citep{kendeigh1982}, \texttt{6531} because the 1978 data point was entered as 7 should have been 47 \citep{minot1986}, and \texttt{6566} because some of the data did not match the graph \citep{heessen1996}.

@@

For the Gompertz model, our weakly-informative priors (Fig.~\ref{fig:priors}) were:
\begin{align*}
b &\sim \mathrm{Uniform}(-1, 2)\\ \lambda &\sim \mathrm{Normal}(0, 10^2)\\
\sigma &\sim \mathrm{Half\mhyphen Cauchy} (0, 2.5)\\ \nu &\sim
\mathrm{Truncated\mhyphen Exponential}(0.01, \mathrm{min.} = 2).
\end{align*}
Our prior on $b$ was uninformative between values of $-1$ and $2$. We would not expect values of $b$ with levels of density dependence as low as $-1$ (very strong inverse density dependence), nor would we generally expect values above $1$. We allowed values of $b$ above $1$ to allow for non-stationary time series of growth rates. The estimates of $b$ were well within these bounds. Our prior on $\lambda$ was very weakly informative within the range of expected values for population growth and is similar to the default priors suggested by \citet{gelman2008d} for intercepts of regression models. Our Half-Cauchy prior on $\sigma$ follows \citet{gelman2006c} and \citet{gelman2008d} and the scale parameter of $2.5$ is based on our expected range of the value in nature from previous studies \citep[e.g.][]{connors2014}. In our testing of a subsample of populations, our parameter estimates were not qualitatively changed by switching to an uninformative uniform prior on $\sigma$, but the weakly informative Half-Cauchy prior substantially sped up chain convergence.

@@

To test if the prior on $\nu$ influenced our estimate of black-swan dynamics, we refit our models with weaker and stronger priors. Our base model used a prior on $\nu$ of Truncated-Exponential(0.02, min.\ = 2). For a weaker prior we used Truncated-Exponential(0.005, min.\ = 2) and for a stronger prior we used Truncated-Exponential(0.02, min.\ = 2) (Fig.~\ref{fig:priors}). Note that the base and weaker priors are relatively flat within the region of $\nu < 20$---the region we are mostly concerned about when categorizing populations as heavy- or thin-tailed.

Our results show that these weaker and stronger priors would have little influence on our conclusions about heavy-tailed dynamics (Fig.~\ref{fig:alt-priors}). When the data are informative about tail behaviour (i.e.\ when there is strong evidence of low $\nu$ values, upper-right of Fig.~\ref{fig:alt-priors}), the prior has little impact on the estimate of $\nu$. When the data are less informative about $\nu$ (i.e.\ when there are no or few tail events and time series are short or noisy), the prior can pull the estimate of $\nu$ towards larger or smaller values (Fig.~\ref{fig:alt-priors}). The vast majority of the populations with Pr$(\nu < 10)$ in the base prior are not altered qualitatively by this range of prior strength.

@@

We considered a version of the Gompertz model in which an autoregressive
parameter was fit to the process noise residuals:
\begin{align*}
x_t &= \lambda + b x_{t-1} + \epsilon_t\\
\epsilon_t &\sim \mathrm{Student\mhyphen t}(\nu, \phi \epsilon_{t-1}, \sigma).
\end{align*}
In addition to the parameters in the original Gompertz model, we estimate an additional parameter $\phi$, which represents the dependence of subsequent residuals. Based on the results of previous analyses with the GPDD \citep[e.g.][]{connors2014} and the chosen priors in previous analyses \citep[e.g.][]{thorson2014a} and to greatly speed up chain convergence when running our model across all populations, we placed a weakly informative prior on $\phi$ that assumed the greatest probability density near zero with the reduced possibility of $\phi$ being near $-1$ or $1$. Specifically, we chose $\phi \sim \mathrm{Truncated\mhyphen Normal}(0, 1, \mathrm{min.} = -1, \mathrm{max.} = 1)$. The MCMC chains did not converge for \modelsNoConvergeAROne\ populations according to our criteria ($\widehat{R} < 1.05, n_\mathrm{eff} > 200$) after 8000 iterations of four chains. This included only \modelsNoConvergeAROneHeavyBase\ populations in which Pr($\nu < 10$) $> 0.5$ categorized them as heavy in the main Gompertz model. We did not include these models in Fig.~\ref{fig:alt}.

@@

We then fit our Gompertz models to the simulated datasets with all parameters (except $\nu$) set near the median values estimated in the GPDD. We repeated this with $50$ and $100$ samples without observation error, $50$ samples with observation error ($\sigma_\mathrm{obs} = 0.2$), and $50$ samples with the same observation error and a Gompertz model that allowed for correctly specified observation error magnitude.

Our results indicate that the Gompertz model can recapture the true value of $\nu$ when the process noise was chosen so that appropriate tail events were present (Fig.~\ref{fig:sim-prob} upper panels). The addition of observation error caused the model to tend to underestimate the degree of heavy-tailedness. Fitting a model with correctly specified observation error did not make substantial improvements to model bias (Fig.~\ref{fig:sim-prob}).

%(Figs~\ref{fig:sim-gompertz} and \ref{fig:sim-gompertz-boxplots}, red and green symbols in the top rows). Likewise, the other Gompertz parameters were estimated without any systematic bias (Figs~\ref{fig:sim-gompertz} and \ref{fig:sim-gompertz-boxplots}, red and green symbols).
%, overestimate the magnitude of process noise, somewhat overestimate $\lambda$, and overestimate density dependence (blue symbols in Figs~\ref{fig:sim-gompertz} and \ref{fig:sim-gompertz-boxplots}). The overestimation of density dependence with observation error is a known phenomenon \citep{knape2012}. Fitting a model with correctly specified observation error made marginal improvements to model bias (purple symbols in Figs~\ref{fig:sim-gompertz} and \ref{fig:sim-gompertz-boxplots}).

When converting the posterior distributions of $\nu$ into Pr($\nu < 10$), the models distinguished heavy and not-heavy tails reasonably well (Fig.~\ref{fig:sim-prob} lower panels). Without observation error, and using a probability of $0.5$ as a threshold, the model correctly classified all simulated systems with normally distributed process noise as not heavy tailed. The model would have miscategorized only one of $40$ simulations at $\nu = 5$ across a sample size of $50$ or $100$ data points (Fig.~\ref{fig:sim-prob}, scenarios 1 and 2 in lower row, second panel from left). The model would have correctly categorized all cases where the process noise was not heavy tailed (Fig.~\ref{fig:sim-prob} bottom-right panel) and all cases where $\nu = 3$ and there was not observation error. With $0.2$ standard deviations of observation error, the model still categorized \obsErrorNuFivePerc\% of cases as heavy tailed when $\nu = 5$ and all but one case when $\nu = 3$. Allowing for observation error made little improvement to the detection of heavy tails. Therefore, we chose to focus on the simpler model without observation error in the main text, particularly given that the true magnitude of observation error was unknown in the empirical data.

@@

\begin{table}
\begin{footnotesize}
\caption{Summary statistics for the filtered Global Population Dynamics Database time series arranged by taxonomic class. Columns are: number of populations, number of taxonomic orders, numbers of species, median time series length, total number of interpolated time steps, total number of substituted zeros, and total number of time steps.}
\smallskip
\begin{tabular}{lrrrrrrrr}
\toprule
\input{../analysis/stat-table.tex}
\label{tab:stats}
\end{tabular}
\end{footnotesize}
\end{table}

@@

\singlespacing
\begin{footnotesize}
\begin{longtable}{>{\RaggedRight}m{1.5cm}>{\RaggedRight}p{4.3cm}>{\RaggedRight}p{0.8cm}>{\RaggedRight}p{1.7cm}>{\RaggedRight}p{1.0cm}>{\RaggedRight}p{3.0cm}>{\RaggedRight}p{1.7cm}>{\RaggedRight}p{1.3cm}}
\caption{All populations with Pr$(\nu < 10) > 0.5$ in the base heavy-tailed Gompertz population dynamics model. Shown are the log abundance time series, population descriptions, Global Population Dynamics Database Main IDs, citation for the data source or separate verification literature, a description of the cause of the black swan events (if known), the probability of heavy tails as calculated by our model, and median estimate of $\nu$ from our model with 90\% quantile credible intervals indicated in parentheses. Red dots on the time series indicate downward black-swan events and blue values indicate upward black-swan events that have a $10 \cdot 10^{-4}$ probability or less of occurring if the population dynamics were explained by a Gompertz model with normally distributed process noise with a standard deviation equal to the scale parameter in the fitted t distribution.}\\
\toprule
\input{../analysis/cause-table.tex}
\label{tab:causes-supp}
\end{longtable}
\end{footnotesize}
\onehalfspacing

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=\textwidth]{../analysis/all-clean-ts-fishes-others.pdf}
\caption{
  (fishes, crustaceans, gastropods, sharks). All filtered time series used in our analysis. The abundances are shown on a log10 vertical axis. Throughout this figure, red dots indicate values that were interpolated and blue dots indicate values that were recorded as zero but were set to the next lowest observed abundance. Numbers before each species name are the GPDD Main ID numbers.
}
\label{fig:all-ts}
\end{center}
\end{figure}

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.8\textwidth]{../analysis/priors-gomp-base.pdf}
\caption{
Probability density of the Bayesian priors for the Gompertz models. From left to right and then top to bottom: (1) per capita growth rate at $\log$(abundance) = $0$: $\lambda \sim \mathrm{Normal}(0, 10^2)$; (2) scale parameter of t-distribution process noise: $\sigma \sim \mathrm{Half\mhyphen Cauchy} (0, 2.5)$; (3) t-distribution degrees of freedom parameter: $\nu \sim \mathrm{Truncated\mhyphen Exponential}(0.01, \mathrm{min.} = 2)$; (4) AR1 correlation coefficient of residuals: $\phi \sim \mathrm{Truncated \mhyphen Normal}(0, 1, \mathrm{min.} = -1, \mathrm{max.} = 1)$. Not shown is $b$, the density dependence parameter: $b \sim \mathrm{Uniform}(-1, 2)$. The $\nu$ panel also shows two alternative priors: a weaker prior $\nu \sim \mathrm{Truncated\mhyphen Exponential}(0.005, \mathrm{min.} = 2)$, and a stronger prior $\nu \sim \mathrm{Truncated\mhyphen Exponential}(0.02, \mathrm{min.} = 2)$. The inset panel shows the same data but with a log-transformed x axis. Note that the base and weaker priors are relatively flat within the region of $\nu < 20$ that we are concerned with.
}
\label{fig:priors}
\end{center}
\end{figure}

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=\textwidth]{../analysis/gomp-comparison.pdf}
\caption{
  Estimates of $\nu$ from alternative models plotted against the base Gompertz
  model estimates of $\nu$. Shown are medians of the posterior (dots) and 50\%
  credible intervals (segments). The diagonal line indicates a one-to-one
  relationship. Different colours indicate various taxonomic classes. The
  grey-shaded regions indicate regions of disagreement if $\nu = 10$ is taken
  as a threshold of heavy-tailed dynamics. The Gompertz observation error model
  assumes a fixed standard deviation of observation error of $0.2$ on a log
  scale.
}
\label{fig:alt}
\end{center}
\end{figure}

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=\textwidth]{../analysis/gomp-prior-comparison.pdf}
\caption{
Estimates of $\nu$ from Gompertz models with alternative priors on $\nu$. Shown are medians of the posterior (dots) and 50\% credible intervals (segments). The diagonal line indicates a one-to-one relationship. Different colours indicate various taxonomic classes. The grey-shaded regions indicate regions of disagreement if $\nu = 10$ is taken as a threshold of heavy-tailed dynamics. The base, weaker, and stronger priors on $\nu$ are illustrated in Fig.~\ref{fig:priors}. In general, the estimates are nearly identical in cases where the data are informative about low values of $\nu$. When the data are less informative about low values of $\nu$, the prior can slightly pull the estimates of $\nu$ towards higher or lower values.
}
\label{fig:alt-priors}
\end{center}
\end{figure}

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.8\textwidth]{../analysis/t-dist-sampling-sim-prior-exp0point01.pdf}
\includegraphics[width=0.8\textwidth]{../analysis/t-dist-sampling-sim-sigma-prior-exp0point01.pdf}
\caption{
  Testing the ability to estimate $\nu$ (top panels) and the scale parameter of
  the process deviations (bottom panels) for a given number of samples (columns)
  drawn from a distribution with a given true $\nu$ value (rows). The red lines
  indicate the true population value. When a small number of samples are drawn
  there may not be samples sufficiently far into the tails to recapture the
  true $\nu$ value; however, heavy tails are still distinguished from normal
  tails in most cases, even with only 25 or 50 samples.
}
\label{fig:sim-nu}
\end{center}
\end{figure}

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=\textwidth]{../analysis/sim-gompertz-median-dist.pdf}
\includegraphics[width=\textwidth]{../analysis/sim-gompertz-p10.pdf}
\caption{
Simulation testing the Gompertz estimation model when the process deviation
draws were chosen so that $\nu$ could be estimated close to the true value
outside the full population model (``effective $\nu$'' within a CV of 0.2 of
specified $\nu$).
Upper panels show the distribution of median $\widehat{\nu}$ across 20 simulation runs.
Lower panels show the distribution of Pr($\nu < 10$) across 20 simulation runs.
We ran the simulations across three population (``true'') $\nu$ values
(3, 5, and $1\cdot 10^9$, i.e.\ approximately normal) and four scenarios: (1) 100 time steps and no
observation error, (2) 50 time steps and no observation error, (3) 50 time
steps and observation error drawn from $\mathrm{Normal} (0, 0.2^2)$ but
ignored, and (4) 50 time steps with observation error in which the quantity
of observation error was assumed known.
Within each scenario the dots represent stochastic draws from the true population
distributions combined with model fits.
}
\label{fig:sim-prob}
\end{center}
\end{figure}

>>>>

\item We removed all populations with uneven sampling intervals, i.e.\ we
  removed populations that did not have a constant difference between the
  ``decimal year begin'' and ``decimal year end'' columns.

@@

\item Finally, we removed populations with the following GPDD Main IDs, which
  we discovered were data entry errors when verifying the populations with
  suspected black swans: \texttt{1207} because the 1957 data point was entered
  as 2 but should have been 27 \citep{kendeigh1982}, \texttt{6531} because the
  1978 data point was entered as 7 but should have been 47 \citep{minot1986},
  and \texttt{6566} because some of the data did not match the graph
  \citep{heessen1996}.

@@

For the Gompertz model, our weakly-informative priors (Fig.~\ref{fig:priors})
were:
\begin{align*}
b &\sim \mathrm{Uniform}(-1, 2)\\ \lambda &\sim \mathrm{Normal}(0, 10^2)\\
\sigma &\sim \mathrm{Half\mhyphen Cauchy} (0, 2.5)\\ \nu &\sim
\mathrm{Truncated\mhyphen Exponential}(0.01, \mathrm{min.} = 2).
\end{align*}
Our prior on $b$ was uninformative between values of $-1$ and $2$. We would
not expect values of $b$ with levels of density dependence as low as $-1$
(very strong inverse density dependence), nor would we generally expect values
above $1$. We allowed values of $b$ above $1$ to allow for non-stationary time
series of growth rates. The estimates of $b$ were well within these bounds.
Our prior on $\lambda$ was very weakly informative within the range of
expected values for population growth and is similar to the default priors
suggested by \citet{gelman2008d} for intercepts of regression models. Our
Half-Cauchy prior on $\sigma$ follows \citet{gelman2006c} and
\citet{gelman2008d} and the specific scale parameter of $2.5$ is based on our
expected range of the value in nature from previous studies
\citep[e.g.][]{connors2014}. In our testing of a subsample of populations, our
parameter estimates were not qualitatively changed by switching to an
uninformative uniform prior on $\sigma$, but the weakly informative
Half-Cauchy prior substantially sped up chain convergence.

@@

To test if the prior on $\nu$ influenced our estimate of black-swan dynamics,
we refit our models with weaker and stronger priors. Our base model used a
prior on $\nu$ of Truncated-Exponential(0.02, min.\ = 2). For a weaker prior
we used Truncated-Exponential(0.005, min.\ = 2) and for a stronger prior we
used Truncated-Exponential(0.02, min.\ = 2) (Fig.~\ref{fig:priors}). Note that
the base and weaker priors are relatively flat within the region of $\nu <
20$, which is the region we are mostly concerned about when categorizing populations
as heavy- or thin-tailed.

Our results show that these weaker and stronger priors would have little
influence on our conclusions about heavy-tailed dynamics
(Fig.~\ref{fig:alt-priors}). When the data are informative about tail
behaviour (i.e.\ when there is strong evidence of low $\nu$ values,
upper-right of Fig.~\ref{fig:alt-priors}), the prior has little impact on the
estimate of $\nu$. When the data are less informative about $\nu$ (i.e.\ when
there are no or few tail events and time series are short or noisy), the prior
can pull the estimate of $\nu$ towards larger or smaller values
(Fig.~\ref{fig:alt-priors}). The vast majority of the populations with Pr$(\nu
< 10)$ in the base prior were not altered qualitatively by this range of prior
strength.

@@

We considered a version of the Gompertz model in which an autoregressive
parameter was fit to the process noise residuals:
\begin{align*}
x_t &= \lambda + b x_{t-1} + \epsilon_t\\
\epsilon_t &\sim \mathrm{Student\mhyphen t}(\nu, \phi \epsilon_{t-1}, \sigma).
\end{align*}
In addition to the parameters in the original Gompertz model, this model
estimates an additional parameter $\phi$, which represents the correlation of
subsequent residuals. Based on the results of previous analyses with the GPDD
\citep[e.g.][]{connors2014} and the chosen priors in previous analyses
\citep[e.g.][]{thorson2014a} and to greatly speed up chain convergence when
running our model across all populations, we placed a weakly informative prior
on $\phi$ that assumed the greatest probability density near zero with the
reduced possibility of $\phi$ being near $-1$ or $1$. Specifically, we chose
$\phi \sim \mathrm{Truncated\mhyphen Normal}(0, 1, \mathrm{min.} = -1,
\mathrm{max.} = 1)$. The MCMC chains did not converge for
\modelsNoConvergeAROne\ populations according to our criteria ($\widehat{R} <
1.05, n_\mathrm{eff} > 200$) after 8000 iterations of four chains. This
  included only \modelsNoConvergeAROneHeavyBase\ populations in which Pr($\nu
  < 10$) $> 0.5$ categorized them as heavy in the main Gompertz model. We did
  not include these models in Fig.~\ref{fig:alt}.

@@

We then fit our Gompertz models to the simulated datasets with all parameters
(except $\nu$) set near the median values estimated in the GPDD. We repeated
this with $50$ and $100$ samples without observation error, $50$ samples with
observation error ($\sigma_\mathrm{obs} = 0.2$), and $50$ samples with the
same observation error and a Gompertz model that allowed for correctly
specified observation error magnitude. Our results indicate that the Gompertz
model can recapture the true value of $\nu$ when the process noise was chosen
so that appropriate tail events were present (Fig.~\ref{fig:sim-prob} upper
panels). The addition of observation error caused the model to tend to
underestimate the degree of heavy-tailedness. Fitting a model with correctly
specified observation error did not make substantial improvements to model
bias (Fig.~\ref{fig:sim-prob}).

%(Figs~\ref{fig:sim-gompertz} and \ref{fig:sim-gompertz-boxplots}, red and
%green symbols in the top rows). Likewise, the other Gompertz parameters were
%estimated without any systematic bias (Figs~\ref{fig:sim-gompertz} and
%\ref{fig:sim-gompertz-boxplots}, red and green symbols). , overestimate the
%magnitude of process noise, somewhat overestimate $\lambda$, and overestimate
%density dependence (blue symbols in Figs~\ref{fig:sim-gompertz} and
%\ref{fig:sim-gompertz-boxplots}). The overestimation of density dependence
%with observation error is a known phenomenon \citep{knape2012}. Fitting a
%model with correctly specified observation error made marginal improvements
%to model bias (purple symbols in Figs~\ref{fig:sim-gompertz} and
%\ref{fig:sim-gompertz-boxplots}).

When converting the posterior distributions of $\nu$ into Pr($\nu < 10$), the
models distinguished heavy and not-heavy tails reasonably well
(Fig.~\ref{fig:sim-prob} lower panels). Without observation error, and using a
probability of $0.5$ as a threshold, the model correctly classified all
simulated systems with normally distributed process noise as not heavy tailed.
The model would have miscategorized only one of $40$ simulations at $\nu = 5$
across simulated populations with $50$ or $100$ time steps
(Fig.~\ref{fig:sim-prob}, scenarios 1 and 2 in lower row, second panel from
left). The model would have correctly categorized all cases where the process
noise was not heavy tailed (Fig.~\ref{fig:sim-prob} bottom-right panel) and
all cases where $\nu = 3$ and there was not observation error. With $0.2$
standard deviations of observation error, the model still categorized
\obsErrorNuFivePerc\% of cases as heavy tailed when $\nu = 5$ and all but one
case when $\nu = 3$. Allowing for observation error made little improvement to
the detection of heavy tails. Therefore, we chose to focus on the simpler
model without observation error in the main text, particularly given that the
true magnitude of observation error was unknown in the empirical data.

@@

\begin{table}
\begin{footnotesize}

\caption{Summary statistics for the filtered Global Population Dynamics
  Database time series arranged by taxonomic class. Columns are: number of
  populations, number of taxonomic orders, numbers of species, median time
  series length, total number of interpolated time steps, total number of
  substituted zeros, and total number of time steps.}

\smallskip
\begin{tabular}{lrrrrrrrr}
\toprule
\input{../analysis/stat-table.tex}
\label{tab:stats}
\end{tabular}
\end{footnotesize}
\end{table}

@@

\singlespacing
\begin{footnotesize}
\begin{longtable}{>{\RaggedRight}m{1.5cm}>{\RaggedRight}p{4.3cm}>{\RaggedRight}p{0.8cm}>{\RaggedRight}p{1.7cm}>{\RaggedRight}p{1.0cm}>{\RaggedRight}p{3.0cm}>{\RaggedRight}p{1.7cm}>{\RaggedRight}p{1.3cm}}

\caption{All populations with Pr$(\nu < 10) > 0.5$ in the base heavy-tailed
  Gompertz population dynamics model. Shown are the log abundance time series,
  population descriptions, Global Population Dynamics Database Main IDs,
  citation for the data source or separate verification literature, a
  description of the cause of the black swan events (if known), the
  probability of heavy tails as calculated by our model, and median estimate
  of $\nu$ from our model with 90\% quantile credible intervals indicated in
  parentheses. Red dots on the time series indicate downward black-swan events
  and blue values indicate upward black-swan events that have a $10 \cdot
  10^{-4}$ probability or less of occurring if the population dynamics were
  explained by a Gompertz model with normally distributed process noise with a
  standard deviation equal to the scale parameter in the fitted t
  distribution.}\\

\toprule
\input{../analysis/cause-table.tex}
\label{tab:causes-supp}
\end{longtable}
\end{footnotesize}
\onehalfspacing

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=\textwidth]{../analysis/all-clean-ts-fishes-others.pdf}

\caption{(fishes, crustaceans, gastropods, sharks). All filtered time series
  used in our analysis. The abundances are shown on a log10 vertical axis.
  Throughout this figure, red dots indicate values that were interpolated and
  blue dots indicate values that were recorded as zero but were set to the
  next lowest observed abundance. Numbers before each species name are the
  GPDD Main ID numbers.}

\label{fig:all-ts}
\end{center}
\end{figure}

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.8\textwidth]{../analysis/priors-gomp-base.pdf}

\caption{ Probability density of the Bayesian priors for the Gompertz models.
  From left to right and then top to bottom: (1) per capita growth rate at
  $\log$(abundance) = $0$: $\lambda \sim \mathrm{Normal}(0, 10^2)$; (2) scale
  parameter of t-distribution process noise: $\sigma \sim \mathrm{Half\mhyphen
    Cauchy} (0, 2.5)$; (3) t-distribution degrees of freedom parameter: $\nu
  \sim \mathrm{Truncated\mhyphen Exponential}(0.01, \mathrm{min.} = 2)$; (4)
  AR1 correlation coefficient of residuals: $\phi \sim \mathrm{Truncated
    \mhyphen Normal}(0, 1, \mathrm{min.} = -1, \mathrm{max.} = 1)$. Not shown
  is $b$, the density dependence parameter: $b \sim \mathrm{Uniform}(-1, 2)$.
  The $\nu$ panel also shows two alternative priors: a weaker prior $\nu \sim
  \mathrm{Truncated\mhyphen Exponential}(0.005, \mathrm{min.} = 2)$, and a
  stronger prior $\nu \sim \mathrm{Truncated\mhyphen Exponential}(0.02,
  \mathrm{min.} = 2)$. The inset panel shows the same data but with a
  log-transformed x axis. Note that the base and weaker priors are relatively
  flat within the region of $\nu < 20$ that we are concerned with. }

\label{fig:priors}
\end{center}
\end{figure}

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=\textwidth]{../analysis/gomp-comparison.pdf}

\caption{Estimates of $\nu$ from alternative models plotted against the base
  Gompertz model estimates of $\nu$. Shown are medians of the posterior (dots)
  and 50\% credible intervals (segments). The diagonal line indicates a
  one-to-one relationship. Different colours indicate various taxonomic
  classes. The grey-shaded regions indicate regions of disagreement if $\nu =
  10$ is taken as a threshold of heavy-tailed dynamics. The Gompertz
  observation error model assumes a fixed standard deviation of observation
  error of $0.2$ on a log scale.}

\label{fig:alt}
\end{center}
\end{figure}

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=\textwidth]{../analysis/gomp-prior-comparison.pdf}

\caption{Estimates of $\nu$ from Gompertz models with alternative priors on
  $\nu$. Shown are medians of the posterior (dots) and 50\% credible intervals
  (segments). The diagonal line indicates a one-to-one relationship. Different
  colours indicate various taxonomic classes. The grey-shaded regions indicate
  regions of disagreement if $\nu = 10$ is taken as a threshold of
  heavy-tailed dynamics. The base, weaker, and stronger priors on $\nu$ are
  illustrated in Fig.~\ref{fig:priors}. In general, the estimates are nearly
  identical in cases where the data are informative about low values of $\nu$.
  When the data are less informative about low values of $\nu$, the prior can
  slightly pull the estimates of $\nu$ towards higher or lower values.}

\label{fig:alt-priors}
\end{center}
\end{figure}

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.8\textwidth]{../analysis/t-dist-sampling-sim-prior-exp0point01.pdf}
\includegraphics[width=0.8\textwidth]{../analysis/t-dist-sampling-sim-sigma-prior-exp0point01.pdf}

\caption{Testing the ability to estimate $\nu$ (top panels) and the scale
  parameter of the process deviations (bottom panels) for a given number of
  samples (columns) drawn from a distribution with a given true $\nu$ value
  (rows). The red lines indicate the true population value. When a small
  number of samples are drawn there may not be samples sufficiently far into
  the tails to recapture the true $\nu$ value; however, heavy tails are still
  distinguished from normal tails in most cases, even with only 25 or 50
  samples.}

\label{fig:sim-nu}
\end{center}
\end{figure}

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=\textwidth]{../analysis/sim-gompertz-median-dist.pdf}
\includegraphics[width=\textwidth]{../analysis/sim-gompertz-p10.pdf}

\caption{ Simulation testing the Gompertz estimation model when the process
  deviation draws were chosen so that $\nu$ could be estimated close to the
  true value outside the full population model (``effective $\nu$'' within a
  CV of 0.2 of specified $\nu$). Upper panels show the distribution of median
  $\widehat{\nu}$ across 20 simulation runs. Lower panels show the
  distribution of Pr($\nu < 10$) across 20 simulation runs. We ran the
  simulations across three population (``true'') $\nu$ values (3, 5, and
  $1\cdot 10^9$, i.e.\ approximately normal) and four scenarios: (1) 100 time
  steps and no observation error, (2) 50 time steps and no observation error,
  (3) 50 time steps and observation error drawn from $\mathrm{Normal} (0,
  0.2^2)$ but ignored, and (4) 50 time steps with observation error in which
  the quantity of observation error was assumed known. Within each scenario
  the dots represent stochastic draws from the true population distributions
  combined with model fits. Underlayed boxplots show the median, interquartile
  range, and $1.5$ times the interquartile range. }

\label{fig:sim-prob}
\end{center}
\end{figure}
